\documentclass[12pt,a4paper,openright,twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage{disi-thesis}
\usepackage{code-lstlistings}
\usepackage{notes}
\usepackage{shortcuts}
\usepackage{csquotes}
\usepackage{acronym}

\school{\unibo}
\programme{Corso di Laurea Magistrale in Ingegneria e Scienze Informatiche}
\title{Design and implementation of a scalable domain specific language foundation for ScaFi with Scala 3}
\author{Luca Deluigi}
\date{\today}
\subject{Paradigmi di Programmazione e Sviluppo}
\supervisor{Prof. Mirko Viroli}
\cosupervisor{Dott. Gianluca Aguzzi}
\morecosupervisor{Dott. Roberto Casadei}
\session{IV}
\academicyear{2022-2023}

% Definition of acronyms
\acrodef{IoT}{Internet of Thing}
\acrodef{VM}{Virtual Machine}
\acrodef{CAS}{Collective Adaptive System}
\acrodef{MAS}{Multi Agent System}
\acrodef{DSL}{Domain Specific Language}
\acrodef{FC}{Field Calculus}
\acrodef{HFC}{Higher-Order Field Calculus}
\acrodef{XC}{Exchange Calculus}
\acrodef{AST}{Abstract Syntax Tree}
\acrodef{ML}{Meta Language}
\acrodef{DOT}{Dependent Object Types}
\acrodef{FP}{Functional Programming}
\acrodef{JS}{JavaScript}
\acrodef{JVM}{Java Virtual Machine}
\acrodef{API}{Application Programming Interface}
\acrodef{SBT}{Simple Build Tool}
\acrodef{MoSCoW}{Must, Should, Could, Won't}
\acrodef{GUI}{Graphical User Interface}
\acrodef{UML}{Unified Modeling Language}

% Definition of math operators
\DeclareMathOperator{\return}{return\,}
\DeclareMathOperator{\send}{\:send\,}
\DeclareMathOperator{\retsend}{retsend\,}

\mainlinespacing{1.241} % line spacing in mainmatter, comment to default (1)

\begin{document}

\frontmatter\frontispiece

\begin{abstract}	
Max 2000 characters, strict.
\end{abstract}

\begin{dedication} % this is optional
Optional. Max a few lines.
\end{dedication}

\begin{acknowledgements} % this is optional
Optional. Max 1 page.
\end{acknowledgements}

%----------------------------------------------------------------------------------------
\tableofcontents   
\listoffigures     % (optional) comment if empty
\lstlistoflistings % (optional) comment if empty
%----------------------------------------------------------------------------------------

\mainmatter

\input{chapters/0-introduction}

\input{chapters/1-state-of-the-art}

\input{chapters/2-background}

\input{chapters/3-analysis}

\input{chapters/4-design}

\input{chapters/5-implementation}

\input{chapters/6-evaluation}

\input{chapters/7-conclusion-and-future-work}

%----------------------------------------------------------------------------------------
% BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\backmatter

\nocite{*} % comment this to only show the referenced entries from the .bib file

\bibliographystyle{ieeetr}
\bibliography{bibliography}

\end{document}
